---
title: "Workshop 1 - Introduction to R and Tidyverse"
author: "Paige Lieurance"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Benefits of coding in R
* Reproducible
* Accurate
* Can use Stack Overflow (online forum) for any R related questions
* Large community
* Resource sharing
  
# Rstudio

**Rstudio is an integrated development environment (IDE) for R. It has 4 panes:**

* Console
* Source (R files)
* Environment, History, Connections, Tutorials
* Files, Plots, Packages, Help, Viewer

# Files
* Script: a text file for storing and running code (comments have a hash-mark before them)
* R Markdown: an interactive file for writing anything from notes to entire manuscripts - contains code "chunks" for storing and running code - can be easily converted to (knitted into) an html file, a word document, or a PDF

# Variables, Vectors, and Functions
* Variable: an object that stores information, i.e. a number, text, or a vector
* Vector: a data structure that contains multiple elements of the same type (class) - using `c()` designates a vector
* Function: an object that takes arguments (aka numbers or vectors) as inputs, performs a task on those arguments, and sometimes has an output - the output of a function can be stored as a variable
  
# Tricks
* Run button (green play button): run all highlighted or current line of code
* #: comment out text, this allows you to type notes in a script - it is considered good practice to explain your code as you write it
* ?function_name: get help on any function
  
```{r, include = FALSE}

# PRACTICE 

print(paste0("The sum of 3 and 2 is ", 3+2))  #add 3+2
mean(c(1,2,3)) #take the mean of 1, 2, and 3
y <- 3+2  #store the sum of 3 and 2 as a variable called y
y   #print y
remove(y)   #remove  y from the environment
y <- c(1,2,3)   #store the vector of 1, 2, and 3 as a variable called y
mean(y)   #Use the function, mean(), to calculate the mean of vector y (to calculate the mean of 1, 2, and 3) and print the answer (aka output of the function)
z <- mean(y) #Store the output from mean(y) in the variable, z
```

# Character strings

```{r, include = FALSE}

# examples of creating character strings 

my_name <- "Paige"
my_name # prints
dogs <- c("rottweiler", "husky", "malamute", "chihuahua", "golden retriever")
dogs # prints 
```

**Exercises**

* Make a new variable containing the number 13
* Make a new variable 1345 times the variable above (doing the calculation in code)
* Make new variables containing the name of you and your friend
* Make a vector containing your name and your friend's name.  Hint: use the variables you created in 3.

```{r, echo = TRUE}
new <- 13
new2 <- new*1345

my_name <- "Paige"
friends_name <- "Anthony"

c(my_name, friends_name) # creates a vector with these 2 names
```

# Directories and R projects
* R works with your computer's file system 
* Use `getwd()` to find out where R thinks we are located 
  
```{r, include = FALSE}
getwd() # get working directory - locates the current project within files
```

* R Project: a file that points R at whatever folder the file is saved in
* File pane: a mini file explorer for R projects

# Packages and the Tidyverse
* Package: a collection of functions - sometimes packages contain multiple packages, or multiple collections of functions, i.e. tidyverse. 
* Install package by using `install.packages("package-name")`
* Update packages by using `update.packages("package-name")`
* Load packages by using `library("package-name")`
* Tidyverse is a useful collection of packages
* Packages pane: see all installed packages - loaded packages will have a tick

```{r, include = FALSE}
#install.packages("tidyverse") # uncomment to install 
```

Loading tidyverse into R: even though it is installed, R cannot use any of the functions from the tidyverse package until it is loaded into the session

```{r, include = FALSE}
library(tidyverse)
```

You must reload packages every session - it is easiest to include a code chunk at the beginning of your R markdown or script and run the code prior to every session  

# CSV files and Data Frames

```{r, include = FALSE}
read_csv("Bats_data.csv") 
```

**Exercises**

* Store Bats_data.csv in a variable, i.e. bats (you'll need the code above and the assign arrow)
* Use the View() function to look at the data

```{r}
bats <- read_csv("Bats_data.csv")

View(bats)
```

``` {r, include = FALSE}
summary(bats)

bats %>% # piping
  View()
```

# Manipulating Data
Each column in a data frame is a vector

* $: extracts a vector (column) from the data frame
  
```{r, include = FALSE}

bats$Date
bats$Activity

```

# Logicals in R
* `==` Equal to
* `<` Less than
* `>` Greater than
* `<=` Less than or equal to 
* `>=` Greater than or equal to
* `|` or
* `&` and
* `!` not

```{r, include = FALSE}
5 == 5
x <- 5
x == 5
x > 5
x >= 5
x > 5 | x >=5
x > 5 & x >= 5
y<- "hello"
y == "hello"
x > 5 | y == "hello"
x > 5 & y == "hello"

```

# Filtering data

```{r, include = FALSE}
filter(bats, bats$Treatment_thinned == "medium-term")
filter(bats, Treatment_thinned == "medium_term")

bats %>% 
  filter(Treatment_thinned == "medium-term") %>% 
  filter(Foraging > 10)

```
 
**Exercises**

Find the maximum Activity observed on 7/01/2013

* Filter by Date for 7/01/2013
* Save the filtered data as a new variable, i.e. filtered_bats
* Use the max() function on the column Activity from your new data frame.

```{r}
filtered_bats <- bats %>% 
  filter(Date == "7/01/2013")

max(filtered_bats$Activity)
```

Use the length() function to find the sample size of all observations on 7/01/2013 and the overall sample size of the bats data set 

``` {r}

n <- length(filtered_bats)

n

overall_n <- length(bats$Site)

overall_n

```

# T test examples

``` {r t test, include = FALSE}

darwin <- read.table(text=
'pair crossfertilized selffertilized
     1      23.5             17.4
     2      12.0             20.4
     3      21.0             20.0
     4      22.0             20.0
     5      19.1             18.4
     6      21.5             18.6
     7      22.1             18.6
     8      20.4             15.3
     9      18.3             16.5
     10     21.6             18.0
     11     23.3             16.3
     12     21.0             18.0
     13     22.1             12.8
     14     23.0             15.5
     15     12.0             18.0
     16     2.0             18.0',header=TRUE)

?t.test

head(darwin)

# mean and sd of the differences
mean <- mean(darwin$crossfertilized-darwin$selffertilized)
sd <- sd(darwin$crossfertilized-darwin$selffertilized) 

#paired t-test on cross/self fertilized data
#have to specify paired = true if paired, as the default is 2-sample
t.test(darwin$crossfertilized,darwin$selffertilized,paired=TRUE) 

#one-sample t test
t.test(darwin$crossfertilized - darwin$selffertilized,mu=0)

#Welch two-sample t test 
t.test(darwin$crossfertilized,darwin$selffertilized)

#Student two-sample t test
t.test(darwin$crossfertilized,darwin$selffertilized,var.equal=TRUE)

#POST-HOC (a posterior) POWER ANALYSIS
?power.t.test
power.t.test(n=16,delta=mean,sd=sd,sig.level=0.05,type='paired')
#12.96% chance that we will get a significant result with the current sample size

#PRE-HOC (a priori) POWER ANALYSIS
power.t.test(n=NULL,delta=mean,sd=sd,sig.level=0.05,type='paired',power=0.9)
#need 215 samples to detect a significant result with the above parameters

```

# Resources

Personal Recommendations

- ["RStudio Primers" Step by step interactive tutorials](https://rstudio.cloud/learn/primers)
- ["R Resources for Beginners"](https://unsw-coders.netlify.app/resources/2021-03-22-beginner-resources/)
- ["Functions"](https://www.stat.berkeley.edu/~statcur/Workshop2/Presentations/functions.pdf)

Creating and manipulating variables

- ["Making new variables" on environmentalcomputing](http://environmentalcomputing.net/making-new-variables/)
- ["Creating R objects" on DataCarpentry](https://datacarpentry.org/R-ecology-lesson/01-intro-to-r.html)
- ["Software carpentry"'s intor to RStudio'](http://swcarpentry.github.io/r-novice-gapminder/01-rstudio-intro/index.html)

A few resources for loading other file types

- [readxl](https://readxl.tidyverse.org/reference/read_excel.html): Reading data direct from Excel
- [datapasta](https://cran.r-project.org/web/packages/datapasta/): when copying and pasting data from the web
- [googlesheets4](https://googlesheets4.tidyverse.org/): for reading data from google sheets

Online courses

- [Environmental Computing](http://environmentalcomputing.net/) 
- [Software carpentry's R novice lesson](http://swcarpentry.github.io/r-novice-gapminder/)
- [Datacarpentry's lesson on R for ecology](https://datacarpentry.org/R-ecology-lesson/)
- [A paper looking at how much R is used in Ecology](https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecs2.2567)
- [The tidyverse web page](https://www.tidyverse.org/)
- [Hadley Wickham's book R for data Science](https://r4ds.had.co.nz/)


Learning to code and need inspiration?

- [Anyone can code](https://www.youtube.com/watch?v=qYZF6oIZtfc&list=PLzdnOPI1iJNe1WmdkMG-Ca8cLQpdEAL7Q)
- [Coding is the new literacy](https://www.youtube.com/watch?v=MwLXrN0Yguk&list=PLzdnOPI1iJNe1WmdkMG-Ca8cLQpdEAL7Q)
- [What most schools don't teach](https://www.youtube.com/watch?v=nKIu9yen5nc&feature=c4-overview-vl&list=PLzdnOPI1iJNe1WmdkMG-Ca8cLQpdEAL7Q)
- when coding saves you time, [from XKCD](http://xkcd.com/1205/)

Why good code matters

- [Why I want to write nice R code](http://nicercode.github.io/blog/2013-04-05-why-nice-code/)
- [Science has a credibility problem](http://www.economist.com/news/leaders/21588069-scientific-research-has-changed-world-now-it-needs-change-itself-how-science-goes-wrong)



