---
title: "Workshop 4 - AusTraits"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- get Austraits data into R from Zenodo/Github
- for my database: filter Austraits by taxon_name and export all trait observations for these species (or a subset)
- add traits from other dataset Threatened_spp_traits_OEH.xls
- final database should contain at least 50 species with a broad range of functional traits and taxonomic families 

DATABASE HEADINGS
species
- minimum 50 species from CPW, WSDRF, Woko NP
- include all diagnostic species for CPW and WSDRF
- any additional species to add re RBG Mount Annan? 

traits
- seed mass
- other quantitative seed traits 
- dormancy type
- taxonomy (family)
- fruit type (pod, fleshy)
- dispersal mode
- growth form

germination data/protocol info
- see links from Katherine from RBG
- also see word doc for additional websites/places to find germ data 
- Kew Botanic Gardens, SA Seedbank, handbook Charlotte had (by genus)

seed accession data from ABG
- availability in seedbank
- date of collection
- location of collection (provenance - lat + long)
- germination ?

community
- CPW
- WSDRF
- Woko NP 

```{r data}
# INSTALLS
install.packages("remotes")
remotes::install_github("traitecoevo/austraits", build_vignettes = FALSE)

# LIBRARIES
library(remotes)
library(austraits)
library(stringr)
library(tidyverse)

# load data into R
austraits <- load_austraits() # loads most recent version of AusTraits into R

```

# EXTRACTING DATA BY TRAIT

## Example - specific leaf area and all leaf traits

```{r example}

# Extract one particular trait and assigning it to an object
sla_data <- extract_trait(austraits, "specific_leaf_area")

# Extracting all leaf associated data and assigning it to an object
traits <- austraits$traits$trait_name %>% unique()  # all possible traits 
leaf_traits <- traits[str_which(traits, "leaf")] # extracting data where "leaf" occurs in the trait_name
leaf_data <- extract_trait(austraits, leaf_traits) 

```

## Other trait data (for database)
* Seed data
* Dispersal data
* Dormancy data
* Fruit data
* Growth data
* Germination data

```{r broad traits}

# Extracting all seed associated data and assigning it to an object
seed_traits <- traits[str_which(traits, "seed")]
seed_data <- extract_trait(austraits, seed_traits)

dispersal_traits <- traits[str_which(traits, "dispersal")]
dispersal_data <- extract_trait(austraits, dispersal_traits)

dormancy_traits <- traits[str_which(traits, "dormancy")]
dormancy_data <- extract_trait(austraits, dormancy_traits)

fruit_traits <- traits[str_which(traits, "fruit")]
fruit_data <- extract_trait(austraits, fruit_traits)

growth_traits <- traits[str_which(traits, "growth")]
growth_data <- extract_trait(austraits, growth_traits)

germination_traits <- traits[str_which(traits, "germination")]
germination_data <- extract_trait(austraits, germination_traits)
```

# Extracting SINGLE PARTICULAR traits (per trait group) and assigning them to objects

```{r single seed traits}
# SEED TRAITS

## seed_length
seed_length_data <- extract_trait(austraits, "seed_length")

## seed_width
seed_width_data <- extract_trait(austraits, "seed_width")

## seed_mass
seed_mass_data <- extract_trait(austraits, "seed_mass")

## seed_shape
seed_shape_data <- extract_trait(austraits, "seed_shape")

```

```{r single dispersal traits}
# DISPERSAL TRAITS

## dispersal_appendage
dispersal_appendage_data <- extract_trait(austraits, "dispersal_appendage")

## dispersal_syndrome
dispersal_syndrome_data <- extract_trait(austraits, "dispersal_syndrome")

```

```{r single dormancy traits}
# DORMANCY TRAITS

## dormancy_type
dormancy_type_data <- extract_trait(austraits, "dormancy_type")

```

```{r single fruit traits}
# FRUIT TRAITS

## fruit_type
fruit_type_data <- extract_trait(austraits, "fruit_type")

## fruit_type_functional
fruit_type_functional_data <- extract_trait(austraits, "fruit_type_functional")

```

```{r single growth traits}
# GROWTH TRAITS 

## plant_growth_form
plant_growth_form_data <- extract_trait(austraits, "plant_growth_form")

## growth_habit
growth_habit_data <- extract_trait(austraits, "growth_habit")

```

```{r single germination traits}
# GERMINATION TRAITS

## germination_treatment
germination_treatment_data <- extract_trait(austraits, "germination_treatment")

## germination
germination_data <- extract_trait(austraits, "germination")

```

# EXTRACTING TRAIT DATA BY TAXA

## All taxa of interest

### Use the filter button at the top of the dataset to view by taxa or by trait

```{r all taxa data}

# all species of interest
all_species_data <- extract_trait(austraits,
                    trait_names = c("seed_length",
                                    "seed_width",
                                    "seed_mass",
                                    "seed_shape",
                                    "dispersal_appendage",
                                    "dispersal_syndrome",
                                    "dormancy_type",
                                    "fruit_type",
                                    "fruit_type_functional",
                                    "plant_growth_form",
                                    "growth_habit",
                                    "germination_treatment",
                                    "germination"),
                    taxon_names = c("Acacia falcata",
                                  "Acacia decurrens",
                                  "Acacia parramattensis",
                                  "Alectryon subcinereus",
                                  "Angophora floribunda",
                                  "Angophora subvelutina",
                                  "Aphanopetalum resinosum",
                                  "Argyrodendron actinophyllum",
                                  "Aristida ramosa",
                                  "Brachychiton acerifolius",
                                  "Brachychiton populneus",
                                  "Breynia oblongifolia",
                                  "Brunoniella australis",
                                  "Bursaria spinosa",
                                  "Casuarina cunninghamia",
                                  "Cayratia clematidea",
                                  "Chloris truncata",
                                  "Chloris ventricosa",
                                  "Cissus antarctica",
                                  "Citronella moorei",
                                  "Clematis glycinoides",
                                  "Clerodendrum tomentosum",
                                  "Corymbia maculata",
                                  "Cymbopogon refractus",
                                  "Daviesia ulicifolia",
                                  "Dendrocnide excelsa",
                                  "Desmodium varians",
                                  "Dichelachne micrantha",
                                  "Dillwynia sieberi",
                                  "Dodonaea viscosa",
                                  "Dysoxylum fraserianum",
                                  "Einadia hastata",
                                  "Einadia nutans",
                                  "Elymus scabrus",
                                  "Eragrostis leptostachya",
                                  "Eucalyptus acmenoides",
                                  "Eucalyptus biturbinata",
                                  "Eucalyptus crebra",
                                  "Eucalyptus eugeniodes",
                                  "Eucalyptus fibrosa",
                                  "Eucalyptus laevopinea",
                                  "Eucalyptus moluccana",
                                  "Eucalyptus polyanthemos",
                                  "Eucalyptus tereticornis",
                                  "Eustrephus latifolius",
                                  "Ficus macrophylla",
                                  "Geranium homeanum",
                                  "Glycine tabacina",
                                  "Hardenbergia violacea",
                                  "Indigofera australis",
                                  "Melaleuca decora",
                                  "Melaleuca styphelioides",
                                  "Melia azedarach",
                                  "Microlaena stipoides",
                                  "Notolaea longifolia",
                                  "Olearia viscidula",
                                  "Oxytes brachypoda",
                                  "Pandorea pandorana",
                                  "Pittosporum revolutum",
                                  "Plectranthus parviflorus",
                                  "Poa labillardieri",
                                  "Polyscias murrayi",
                                  "Rubus parvifolius",
                                  "Solanum prinophyllum",
                                  "Stephania japonica",
                                  "Syncarpia glomulifera",
                                  "Synoum glandulosum",
                                  "Themeda triandra",
                                  "Toona ciliata",
                                  "Urtica incisa",
                                  "Wahlenbergia gracilis")) 

View(all_species_data[["traits"]])
```

# Creating additional (smaller) datasets of quantitative traits to pull mean values for each species

# SEED MASS OVERALL

```{r seed_mass dataset}

# create new dataset of just seed_mass
all_species_mass <- extract_trait(austraits,
                    trait_names = "seed_mass",
                    taxon_names = c("Acacia falcata",
                                  "Acacia decurrens",
                                  "Acacia parramattensis",
                                  "Alectryon subcinereus",
                                  "Angophora floribunda",
                                  "Angophora subvelutina",
                                  "Aphanopetalum resinosum",
                                  "Argyrodendron actinophyllum",
                                  "Aristida ramosa",
                                  "Brachychiton acerifolius",
                                  "Brachychiton populneus",
                                  "Breynia oblongifolia",
                                  "Brunoniella australis",
                                  "Bursaria spinosa",
                                  "Casuarina cunninghamia",
                                  "Cayratia clematidea",
                                  "Chloris truncata",
                                  "Chloris ventricosa",
                                  "Cissus antarctica",
                                  "Citronella moorei",
                                  "Clematis glycinoides",
                                  "Clerodendrum tomentosum",
                                  "Corymbia maculata",
                                  "Cymbopogon refractus",
                                  "Daviesia ulicifolia",
                                  "Dendrocnide excelsa",
                                  "Desmodium varians",
                                  "Dichelachne micrantha",
                                  "Dillwynia sieberi",
                                  "Dodonaea viscosa",
                                  "Dysoxylum fraserianum",
                                  "Einadia hastata",
                                  "Einadia nutans",
                                  "Elymus scabrus",
                                  "Eragrostis leptostachya",
                                  "Eucalyptus acmenoides",
                                  "Eucalyptus biturbinata",
                                  "Eucalyptus crebra",
                                  "Eucalyptus eugeniodes",
                                  "Eucalyptus fibrosa",
                                  "Eucalyptus laevopinea",
                                  "Eucalyptus moluccana",
                                  "Eucalyptus polyanthemos",
                                  "Eucalyptus tereticornis",
                                  "Eustrephus latifolius",
                                  "Ficus macrophylla",
                                  "Geranium homeanum",
                                  "Glycine tabacina",
                                  "Hardenbergia violacea",
                                  "Indigofera australis",
                                  "Melaleuca decora",
                                  "Melaleuca styphelioides",
                                  "Melia azedarach",
                                  "Microlaena stipoides",
                                  "Notolaea longifolia",
                                  "Olearia viscidula",
                                  "Oxytes brachypoda",
                                  "Pandorea pandorana",
                                  "Pittosporum revolutum",
                                  "Plectranthus parviflorus",
                                  "Poa labillardieri",
                                  "Polyscias murrayi",
                                  "Rubus parvifolius",
                                  "Solanum prinophyllum",
                                  "Stephania japonica",
                                  "Syncarpia glomulifera",
                                  "Synoum glandulosum",
                                  "Themeda triandra",
                                  "Toona ciliata",
                                  "Urtica incisa",
                                  "Wahlenbergia gracilis"))

View(all_species_mass[["traits"]])

# set seed_mass as a numeric variable 
all_species_mass$value <- as.numeric(all_species_mass$value)

# can use if making a for loop
unique((all_species_mass[["traits"]])$taxon_name)
```

# MEAN SEED MASS PER SPECIES

``` {r mean seed_mass per species}

(mean_seed_mass_Acacia_decurrens <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Acacia decurrens" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Acacia_falcata <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Acacia falcata" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Acacia_implexa <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Acacia implexa" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Acacia_parramattensis <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Acacia parramattensis" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Alectryon_subcinereus <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Alectryon subcinereus" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Angophora_floribunda <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Angophora floribunda" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Angophora_subvelutina <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Angophora subvelutina" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Aphanopetalum_resinosum <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Aphanopetalum resinosum" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Argyrodendron_actinophyllum <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Argyrodendron actinophyllum" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Aristida_ramosa <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Aristida ramosa" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Brachychiton_acerifolius <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Brachychiton acerifolius" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Brachychiton_populneus <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Brachychiton populneus" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Breynia_oblongifolia <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Breynia oblongifolia" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Brunoniella_australis <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Brunoniella australis" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Bursaria_spinosa <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Bursaria spinosa" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Casuarina_cunninghamia <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Casuarina cunninghamia" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Cayratia_clematidea <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Cayratia clematidea" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Chloris_truncata <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Chloris truncata" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Chloris_ventricosa <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Chloris ventricosa" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Cissus_antarctica <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Cissus antarctica" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Citronella_moorei <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Citronella moorei" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Clematis_glycinoides <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Clematis glycinoides" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Clerodendrum_tomentosum <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Clerodendrum tomentosum" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Corymbia_maculata <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Corymbia maculata" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Cymbopogon_refractus <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Cymbopogon refractus" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Daviesia_ulicifolia <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Daviesia ulicifolia" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Dendrocnide_excelsa <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Dendrocnide excelsa" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Desmodium_varians <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Desmodium varians" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Dichelachne_micrantha <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Dichelachne micrantha" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Dillwynia_sieberi <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Dillwynia sieberi" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Dodonaea_viscosa <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Dodonaea viscosa" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Dysoxylum_fraserianum <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Dysoxylum fraserianum" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Einadia_hastata <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Einadia hastata" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Einadia_nutans <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Einadia nutans" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Elymus_scabrus <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Elymus scabrus" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eragrostis_leptostachya <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eragrostis leptostachya" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eucalyptus_acmenoides <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eucalyptus acmenoides" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eucalyptus_biturbinata <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eucalyptus biturbinata" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eucalyptus_crebra <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eucalyptus crebra" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eucalyptus_eugeniodes <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eucalyptus eugeniodes" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eucalyptus_fibrosa <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eucalyptus fibrosa" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eucalyptus_laevopinea <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eucalyptus laevopinea" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eucalyptus_moluccana <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eucalyptus moluccana" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eucalyptus_polyanthemos <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eucalyptus polyanthemos" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eucalyptus_tereticornis <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eucalyptus tereticornis" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Eustrephus_latifolius <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Eustrephus latifolius" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Ficus_macrophylla <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Ficus macrophylla" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Geranium_homeanum <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Geranium homeanum" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Glycine_tabacina <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Glycine tabacina" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Hardenbergia_violacea <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Hardenbergia violacea" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Indigofera_australis <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Indigofera australis" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Melaleuca_decora <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Melaleuca decora" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Melaleuca_styphelioides <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Melaleuca styphelioides" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Melia_azedarach <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Melia azedarach" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Microlaena_stipoides <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Microlaena stipoides" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Notolaea_longifolia <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Notolaea longifolia" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Olearia_viscidula <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Olearia viscidula" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Oxytes_brachypoda <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Oxytes brachypoda" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Pandorea_pandorana <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Pandorea pandorana" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Pittosporum_revolutum <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Pittosporum revolutum" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Plectranthus_parviflorus <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Plectranthus parviflorus" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Poa_labillardieri <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Poa labillardieri" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Polyscias_murrayi <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Polyscias murrayi" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Rubus_parvifolius <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Rubus parvifolius" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Solanum_prinophyllum <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Solanum prinophyllum" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Stephania_japonica <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Stephania japonica" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Syncarpia_glomulifera <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Syncarpia glomulifera" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Synoum_glandulosum <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Synoum glandulosum" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Themeda_triandra <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Themeda triandra" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Toona_ciliata <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Toona ciliata" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Urtica_incisa <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Urtica incisa" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

(mean_seed_mass_Wahlenbergia_gracilis <- mean(all_species_mass[["traits"]]$value[all_species_mass[["traits"]]$taxon_name == "Wahlenbergia gracilis" & all_species_mass[["traits"]]$trait_name == "seed_mass"]))

```

# SEED LENGTH OVERALL

```{r seed_length dataset}

# create new dataset of just seed_length
all_species_length <- extract_trait(austraits,
                    trait_names = "seed_length",
                    taxon_names = c("Acacia falcata",
                                  "Acacia decurrens",
                                  "Acacia parramattensis",
                                  "Alectryon subcinereus",
                                  "Angophora floribunda",
                                  "Angophora subvelutina",
                                  "Aphanopetalum resinosum",
                                  "Argyrodendron actinophyllum",
                                  "Aristida ramosa",
                                  "Brachychiton acerifolius",
                                  "Brachychiton populneus",
                                  "Breynia oblongifolia",
                                  "Brunoniella australis",
                                  "Bursaria spinosa",
                                  "Casuarina cunninghamia",
                                  "Cayratia clematidea",
                                  "Chloris truncata",
                                  "Chloris ventricosa",
                                  "Cissus antarctica",
                                  "Citronella moorei",
                                  "Clematis glycinoides",
                                  "Clerodendrum tomentosum",
                                  "Corymbia maculata",
                                  "Cymbopogon refractus",
                                  "Daviesia ulicifolia",
                                  "Dendrocnide excelsa",
                                  "Desmodium varians",
                                  "Dichelachne micrantha",
                                  "Dillwynia sieberi",
                                  "Dodonaea viscosa",
                                  "Dysoxylum fraserianum",
                                  "Einadia hastata",
                                  "Einadia nutans",
                                  "Elymus scabrus",
                                  "Eragrostis leptostachya",
                                  "Eucalyptus acmenoides",
                                  "Eucalyptus biturbinata",
                                  "Eucalyptus crebra",
                                  "Eucalyptus eugeniodes",
                                  "Eucalyptus fibrosa",
                                  "Eucalyptus laevopinea",
                                  "Eucalyptus moluccana",
                                  "Eucalyptus polyanthemos",
                                  "Eucalyptus tereticornis",
                                  "Eustrephus latifolius",
                                  "Ficus macrophylla",
                                  "Geranium homeanum",
                                  "Glycine tabacina",
                                  "Hardenbergia violacea",
                                  "Indigofera australis",
                                  "Melaleuca decora",
                                  "Melaleuca styphelioides",
                                  "Melia azedarach",
                                  "Microlaena stipoides",
                                  "Notolaea longifolia",
                                  "Olearia viscidula",
                                  "Oxytes brachypoda",
                                  "Pandorea pandorana",
                                  "Pittosporum revolutum",
                                  "Plectranthus parviflorus",
                                  "Poa labillardieri",
                                  "Polyscias murrayi",
                                  "Rubus parvifolius",
                                  "Solanum prinophyllum",
                                  "Stephania japonica",
                                  "Syncarpia glomulifera",
                                  "Synoum glandulosum",
                                  "Themeda triandra",
                                  "Toona ciliata",
                                  "Urtica incisa",
                                  "Wahlenbergia gracilis")) 

View(all_species_length[["traits"]])

# set seed_length as a numeric variable 
all_species_length$value <- as.numeric(all_species_length$value)

```

# MEAN SEED LENGTH PER SPECIES

``` {r mean seed_length per species}

(mean_seed_length_Acacia_decurrens <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Acacia decurrens" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Acacia_falcata <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Acacia falcata" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Acacia_implexa <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Acacia implexa" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Acacia_parramattensis <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Acacia parramattensis" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Alectryon_subcinereus <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Alectryon subcinereus" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Angophora_floribunda <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Angophora floribunda" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Angophora_subvelutina <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Angophora subvelutina" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Aphanopetalum_resinosum <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Aphanopetalum resinosum" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Argyrodendron_actinophyllum <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Argyrodendron actinophyllum" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Aristida_ramosa <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Aristida ramosa" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Brachychiton_acerifolius <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Brachychiton acerifolius" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Brachychiton_populneus <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Brachychiton populneus" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Breynia_oblongifolia <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Breynia oblongifolia" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Brunoniella_australis <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Brunoniella australis" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Bursaria_spinosa <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Bursaria spinosa" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Casuarina_cunninghamia <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Casuarina cunninghamia" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Cayratia_clematidea <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Cayratia clematidea" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Chloris_truncata <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Chloris truncata" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Chloris_ventricosa <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Chloris ventricosa" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Cissus_antarctica <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Cissus antarctica" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Citronella_moorei <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Citronella moorei" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Clematis_glycinoides <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Clematis glycinoides" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Clerodendrum_tomentosum <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Clerodendrum tomentosum" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Corymbia_maculata <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Corymbia maculata" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Cymbopogon_refractus <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Cymbopogon refractus" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Daviesia_ulicifolia <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Daviesia ulicifolia" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Dendrocnide_excelsa <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Dendrocnide excelsa" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Desmodium_varians <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Desmodium varians" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Dichelachne_micrantha <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Dichelachne micrantha" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Dillwynia_sieberi <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Dillwynia sieberi" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Dodonaea_viscosa <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Dodonaea viscosa" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Dysoxylum_fraserianum <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Dysoxylum fraserianum" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Einadia_hastata <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Einadia hastata" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Einadia_nutans <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Einadia nutans" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Elymus_scabrus <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Elymus scabrus" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eragrostis_leptostachya <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eragrostis leptostachya" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eucalyptus_acmenoides <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eucalyptus acmenoides" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eucalyptus_biturbinata <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eucalyptus biturbinata" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eucalyptus_crebra <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eucalyptus crebra" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eucalyptus_eugeniodes <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eucalyptus eugeniodes" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eucalyptus_fibrosa <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eucalyptus fibrosa" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eucalyptus_laevopinea <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eucalyptus laevopinea" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eucalyptus_moluccana <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eucalyptus moluccana" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eucalyptus_polyanthemos <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eucalyptus polyanthemos" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eucalyptus_tereticornis <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eucalyptus tereticornis" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Eustrephus_latifolius <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Eustrephus latifolius" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Ficus_macrophylla <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Ficus macrophylla" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Geranium_homeanum <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Geranium homeanum" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Glycine_tabacina <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Glycine tabacina" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Hardenbergia_violacea <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Hardenbergia violacea" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Indigofera_australis <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Indigofera australis" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Melaleuca_decora <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Melaleuca decora" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Melaleuca_styphelioides <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Melaleuca styphelioides" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Melia_azedarach <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Melia azedarach" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Microlaena_stipoides <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Microlaena stipoides" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Notolaea_longifolia <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Notolaea longifolia" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Olearia_viscidula <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Olearia viscidula" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Oxytes_brachypoda <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Oxytes brachypoda" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Pandorea_pandorana <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Pandorea pandorana" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Pittosporum_revolutum <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Pittosporum revolutum" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Plectranthus_parviflorus <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Plectranthus parviflorus" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Poa_labillardieri <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Poa labillardieri" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Polyscias_murrayi <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Polyscias murrayi" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Rubus_parvifolius <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Rubus parvifolius" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Solanum_prinophyllum <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Solanum prinophyllum" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Stephania_japonica <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Stephania japonica" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Syncarpia_glomulifera <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Syncarpia glomulifera" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Synoum_glandulosum <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Synoum glandulosum" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Themeda_triandra <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Themeda triandra" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Toona_ciliata <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Toona ciliata" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Urtica_incisa <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Urtica incisa" & all_species_length[["traits"]]$trait_name == "seed_length"]))

(mean_seed_length_Wahlenbergia_gracilis <- mean(all_species_length[["traits"]]$value[all_species_length[["traits"]]$taxon_name == "Wahlenbergia gracilis" & all_species_length[["traits"]]$trait_name == "seed_length"]))

```

# SEED WIDTH OVERALL

```{r seed_width dataset}

# create new dataset of just seed_width
all_species_width <- extract_trait(austraits,
                    trait_names = "seed_width",
                    taxon_names = c("Acacia falcata",
                                  "Acacia decurrens",
                                  "Acacia parramattensis",
                                  "Alectryon subcinereus",
                                  "Angophora floribunda",
                                  "Angophora subvelutina",
                                  "Aphanopetalum resinosum",
                                  "Argyrodendron actinophyllum",
                                  "Aristida ramosa",
                                  "Brachychiton acerifolius",
                                  "Brachychiton populneus",
                                  "Breynia oblongifolia",
                                  "Brunoniella australis",
                                  "Bursaria spinosa",
                                  "Casuarina cunninghamia",
                                  "Cayratia clematidea",
                                  "Chloris truncata",
                                  "Chloris ventricosa",
                                  "Cissus antarctica",
                                  "Citronella moorei",
                                  "Clematis glycinoides",
                                  "Clerodendrum tomentosum",
                                  "Corymbia maculata",
                                  "Cymbopogon refractus",
                                  "Daviesia ulicifolia",
                                  "Dendrocnide excelsa",
                                  "Desmodium varians",
                                  "Dichelachne micrantha",
                                  "Dillwynia sieberi",
                                  "Dodonaea viscosa",
                                  "Dysoxylum fraserianum",
                                  "Einadia hastata",
                                  "Einadia nutans",
                                  "Elymus scabrus",
                                  "Eragrostis leptostachya",
                                  "Eucalyptus acmenoides",
                                  "Eucalyptus biturbinata",
                                  "Eucalyptus crebra",
                                  "Eucalyptus eugeniodes",
                                  "Eucalyptus fibrosa",
                                  "Eucalyptus laevopinea",
                                  "Eucalyptus moluccana",
                                  "Eucalyptus polyanthemos",
                                  "Eucalyptus tereticornis",
                                  "Eustrephus latifolius",
                                  "Ficus macrophylla",
                                  "Geranium homeanum",
                                  "Glycine tabacina",
                                  "Hardenbergia violacea",
                                  "Indigofera australis",
                                  "Melaleuca decora",
                                  "Melaleuca styphelioides",
                                  "Melia azedarach",
                                  "Microlaena stipoides",
                                  "Notolaea longifolia",
                                  "Olearia viscidula",
                                  "Oxytes brachypoda",
                                  "Pandorea pandorana",
                                  "Pittosporum revolutum",
                                  "Plectranthus parviflorus",
                                  "Poa labillardieri",
                                  "Polyscias murrayi",
                                  "Rubus parvifolius",
                                  "Solanum prinophyllum",
                                  "Stephania japonica",
                                  "Syncarpia glomulifera",
                                  "Synoum glandulosum",
                                  "Themeda triandra",
                                  "Toona ciliata",
                                  "Urtica incisa",
                                  "Wahlenbergia gracilis")) 

View(all_species_width[["traits"]])

# set seed_length as a numeric variable 
all_species_width$value <- as.numeric(all_species_width$value)

```

# MEAN SEED WIDTH PER SPECIES

``` {r mean seed_width per species}

(mean_seed_width_Acacia_decurrens <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Acacia decurrens" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Acacia_falcata <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Acacia falcata" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Acacia_implexa <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Acacia implexa" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Acacia_parramattensis <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Acacia parramattensis" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Alectryon_subcinereus <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Alectryon subcinereus" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Angophora_floribunda <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Angophora floribunda" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Angophora_subvelutina <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Angophora subvelutina" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Aphanopetalum_resinosum <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Aphanopetalum resinosum" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Argyrodendron_actinophyllum <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Argyrodendron actinophyllum" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Aristida_ramosa <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Aristida ramosa" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Brachychiton_acerifolius <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Brachychiton acerifolius" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Brachychiton_populneus <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Brachychiton populneus" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Breynia_oblongifolia <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Breynia oblongifolia" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Brunoniella_australis <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Brunoniella australis" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Bursaria_spinosa <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Bursaria spinosa" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Casuarina_cunninghamia <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Casuarina cunninghamia" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Cayratia_clematidea <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Cayratia clematidea" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Chloris_truncata <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Chloris truncata" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Chloris_ventricosa <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Chloris ventricosa" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Cissus_antarctica <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Cissus antarctica" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Citronella_moorei <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Citronella moorei" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Clematis_glycinoides <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Clematis glycinoides" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Clerodendrum_tomentosum <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Clerodendrum tomentosum" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Corymbia_maculata <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Corymbia maculata" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Cymbopogon_refractus <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Cymbopogon refractus" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Daviesia_ulicifolia <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Daviesia ulicifolia" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Dendrocnide_excelsa <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Dendrocnide excelsa" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Desmodium_varians <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Desmodium varians" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Dichelachne_micrantha <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Dichelachne micrantha" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Dillwynia_sieberi <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Dillwynia sieberi" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Dodonaea_viscosa <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Dodonaea viscosa" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Dysoxylum_fraserianum <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Dysoxylum fraserianum" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Einadia_hastata <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Einadia hastata" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Einadia_nutans <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Einadia nutans" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Elymus_scabrus <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Elymus scabrus" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eragrostis_leptostachya <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eragrostis leptostachya" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eucalyptus_acmenoides <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eucalyptus acmenoides" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eucalyptus_biturbinata <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eucalyptus biturbinata" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eucalyptus_crebra <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eucalyptus crebra" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eucalyptus_eugeniodes <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eucalyptus eugeniodes" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eucalyptus_fibrosa <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eucalyptus fibrosa" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eucalyptus_laevopinea <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eucalyptus laevopinea" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eucalyptus_moluccana <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eucalyptus moluccana" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eucalyptus_polyanthemos <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eucalyptus polyanthemos" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eucalyptus_tereticornis <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eucalyptus tereticornis" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Eustrephus_latifolius <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Eustrephus latifolius" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Ficus_macrophylla <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Ficus macrophylla" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Geranium_homeanum <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Geranium homeanum" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Glycine_tabacina <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Glycine tabacina" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Hardenbergia_violacea <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Hardenbergia violacea" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Indigofera_australis <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Indigofera australis" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Melaleuca_decora <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Melaleuca decora" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Melaleuca_styphelioides <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Melaleuca styphelioides" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Melia_azedarach <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Melia azedarach" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Microlaena_stipoides <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Microlaena stipoides" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Notolaea_longifolia <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Notolaea longifolia" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Olearia_viscidula <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Olearia viscidula" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Oxytes_brachypoda <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Oxytes brachypoda" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Pandorea_pandorana <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Pandorea pandorana" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Pittosporum_revolutum <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Pittosporum revolutum" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Plectranthus_parviflorus <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Plectranthus parviflorus" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Poa_labillardieri <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Poa labillardieri" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Polyscias_murrayi <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Polyscias murrayi" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Rubus_parvifolius <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Rubus parvifolius" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Solanum_prinophyllum <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Solanum prinophyllum" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Stephania_japonica <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Stephania japonica" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Syncarpia_glomulifera <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Syncarpia glomulifera" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Synoum_glandulosum <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Synoum glandulosum" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Themeda_triandra <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Themeda triandra" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Toona_ciliata <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Toona ciliata" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Urtica_incisa <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Urtica incisa" & all_species_width[["traits"]]$trait_name == "seed_width"]))

(mean_seed_width_Wahlenbergia_gracilis <- mean(all_species_width[["traits"]]$value[all_species_width[["traits"]]$taxon_name == "Wahlenbergia gracilis" & all_species_width[["traits"]]$trait_name == "seed_width"]))

```

# Extra: creating for loop functions to determine multiple means 
## Can be used in the future if mean trait values are needed for many species

``` {r for loops}

# for loop code to determine multiple means (can use for the future)
data <- mtcars

df_names <- c("Cylinders", "Mean HP", "Mean Weight") # create titles for dataframe

car_means <- NULL # create empty dataframe

for(i in unique(data$cyl)){ # calculates and selects each unique value in 'cyl'
  mean_cyl <- mean(data$hp[data$cyl == i], na.rm = TRUE) # calculcates the mean HP of each 'cyl'
  mean_wt <- mean(data$wt[data$cyl == i], na.rm = TRUE) # calculcates the mean weight of each 'cyl'
  car_means <- rbind(car_means,c(i, mean_cyl, mean_wt)) # binds new entry in new row per 'cyl'
}

colnames(car_means) <- df_names # apply names to dataframe

car_means # display dataframe

```

# Meeting with Ruby
## Tips for tackling AusTraits coming into next year

- instead of manually pulling means of quantitative data, try the aggregate function

- even better yet, use the pipe function in Tidyverse:
        - filter by species of interest first, then group by trait, pipe down,             and mutate to find the mean per species
        
- also might be useful to pull all available trait data per species (just to see what is available and any other potential correlates of establishment success)

- Ruby finds pulling different tables per trait and then joining them is useful for collating data of interest 
